---
title: "2018_TOC"
author: "Samantha Chen"
date: "11/11/2020"
output: github_document
---

# Intro

This document will show how **individual bottle** TOC/DOC data from 2018 remineralization experiments were processed, QC'd, and analyzed.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

```{r}
excel_sheets("~/GitHub EEMB 144L/144l_students/Input_Data/week4/144L_2018_Exp_TOC.xlsx")

metadata <- read_excel("~/GitHub EEMB 144L/144l_students/Input_Data/week4/144L_2018_Exp_TOC.xlsx", sheet = "Metadata")

data <- read_excel("~/GitHub EEMB 144L/144l_students/Input_Data/week4/144L_2018_Exp_TOC.xlsx", sheet = "Data")

joined_data <- left_join(metadata, data)
```
```{r}
toc <- joined_data %>% 
  mutate(Datetime = ymd_hm(Datetime)) %>% 
  group_by(Bottle, Treatment) %>% 
  mutate(interv = interval(first(Datetime), Datetime),
         hours = as.numeric(interv)/3600,
         days = hours/24) %>% 
  ungroup() %>%
  rename(sd_TOC = TOC_sd,
         sd_PTOC = PTOC_sd) %>% 
  select(Experiment:Datetime, DNA_SampleID:days)
```

```{r}
pivot_toc <- toc %>% 
  select(Experiment, Location, Bottle, Treatment, days, TOC, PTOC) %>% 
  pivot_longer(TOC:PTOC, names_to = "sample", values_to = "value") #expand rows, decrease columns

pivot_toc_sd <- toc %>% 
  select(Experiment, Location, Bottle, Treatment, days, sd_TOC, sd_PTOC) %>% 
  pivot_longer(sd_TOC:sd_PTOC, names_to = "sample", names_prefix = "sd_", values_to = "sd")

pivoted <- left_join(pivot_toc, pivot_toc_sd) %>% 
  mutate(sample = ifelse(sample == "TOC", "Bottle", "Vial"))

view(pivoted)
```


# TOC Plot

```{r fig.height=10, fig.width=8}
custom.colors <- c("Control" = "#2c7fb8", "Ash Leachate" = "#7fcdbb", "Mud Leachate" = "#756bb1", "Glucose_Nitrate_Phosphate" = "#fc8d62", "Bottle" = "#c51b8a", "Vial" = "#1f78b4")
levels <- c("Control", "Ash Leachate", "Mud Leachate", "Glucose_Nitrate_Phosphate", "Bottle", "Vial")

pivoted %>% 
  drop_na(value) %>% 
  mutate(Treatment = factor(Treatment, levels = levels),
        sample = factor(sample, levels = levels)) %>% 
  ggplot(aes(x = days, y = value, group = interaction(Treatment, Bottle))) +
  geom_errorbar(aes(ymin = value - sd, ymax = value + sd, color = sample), width = 0.4) +
  geom_point(aes(fill = sample), size = 3, shape = 21, alpha = 0.7) +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  labs(x = "Days", y = expression("Total Organic Carbon µmol C L"^-1), color = "", fill = "") +
  theme_bw() +
  facet_grid(rows = "Treatment", scales = "free") +
  guides(color = F)
  
```
Ash and GNP follows a very similar trend in which TOC decreases as days increases. You don't see a dramatic decrease in TOC in Mud and Control compared to Ash and GNP.

The analytical detection limit for our TOC analysis is ~1.5 µmol C L ^-1.

# TOC v. PTOC (Bottle v. Vial Samples)

```{r message=FALSE, warning=FALSE}
# install.packages("lmodel2")
library(lmodel2)

reg.data <- toc %>% 
  drop_na(TOC) %>% 
  filter(Timepoint > 0)

reg <- lmodel2(PTOC ~ TOC, data = reg.data, nperm = 99)

reg
```
```{r}
intercept <- reg$regression.results[3,2]
slope <- reg$regression.results[3,3]

two_int <- reg$confidence.intervals[3,2]
two_slope <- reg$confidence.intervals[3,4]
nine_int <- reg$confidence.intervals[3,3]
nine_slope <- reg$confidence.intervals[3,5]
```
```{r fig.height=4, fig.width=7}
reg.data %>% 
  ggplot(aes(x = TOC, y = PTOC)) +
  geom_errorbar(aes(ymin = PTOC - sd_PTOC, ymax = PTOC + sd_PTOC), width = 0.05) +
  # geom_errorbarh(aes(xmin = TOC - sd_TOC, xmax = PTOC + sd_PTOC), width = 0.05) +
  geom_point(fill = "white", shape = 21, size = 4, alpha = 0.7) +
  geom_abline(intercept = intercept, slope = slope, linetype = 2, size = 1) +
  geom_abline(intercept = two_int, slope = two_slope, color = "black", linetype = 3, size = 1) +
  geom_abline(intercept = nine_int, slope = nine_slope, color = "black", linetype = 3, size = 1) +
  labs(x = expression("Bottle TOC, µmol C L"^-1), y = expression("Vial TOC, µmol C L"^-1)) +
  theme_bw() +
  annotate( geom = "text", label = expression(atop("y = 1.03x - 2.12", paste("r"^2," = 0.97, ", italic("p "), "= 0.01"))), x = 90, y = 85, size = 3)
```

This regression indicates a positive correlation between Bottle TOC and Vial TOC. Slope = 1.03 µmol C L ^-1 and intercept = -2.12. R value = 0.97 indicating a strong linear relationship. P value = 0.01 which is <0.05 indicating a statistically significant correlation.

```{r}
bc <- read_rds("~/GitHub EEMB 144L/144l_students/Output_Data/week 4/2018_BactAbund_files/2018_BactAbund.rds") %>% 
  select(-c(hours, days))

merge <- left_join(toc, bc) %>% 
  select(Experiment:days, TOC:sd_PTOC, cells:diff_ln_cells, bc, ave_bc, sd_bc, everything(), -c(contains("Sample"))) %>% 
  group_by(Experiment, Treatment, Bottle) %>% 
  fill(start_exp:sd_lag) %>% 
  ungroup()

glimpse(merge)
  
```

# Interpolation

```{r message=FALSE, warning=FALSE}
#install.packages("zoo")
library(zoo)
```

```{r}
to_interpolate <- merge %>% 
  select(Experiment, Bottle, Timepoint, days, PTOC, bc) %>% 
  group_by(Experiment, Bottle)

list <- to_interpolate %>% 
  group_split()

keys <- to_interpolate %>% 
  group_keys() %>% 
  mutate(key = paste(Experiment, Bottle)) %>% 
  ungroup()

names(list) <- keys$key
```

```{r}
interp.func <- function(x) {
  y <- zoo(x, order.by = x$days)
  interp_toc <- round(as.numeric(na.approx(y$PTOC, na.rm = 
                                              F)), 1)
  interp_bc <- round(as.numeric(na.approx(y$bc, na.rm = F)), 1)
  z <- cbind(y, interp_toc, interp_bc)
  as_tibble(z)
}
```
```{r}
Interpolated <- lapply(list, interp.func) %>% 
  plyr::ldply(., as.data.frame) %>% 
  select(-c(.id, PTOC, bc, days)) %>% 
  mutate_at(vars(Timepoint:interp_bc), as.numeric) %>% 
  left_join(merge, .)
  
```

# Estimate DOC, Bioavailability, BGEs

```{r}
doc <- Interpolated %>% 
  mutate(doc = interp_toc - interp_bc) %>% 
  group_by(Experiment, Bottle) %>% 
  mutate(bioav_doc = (first(doc) - last(na.omit(doc))) / first(doc), 
         doc_end_exp = ifelse(Timepoint == end_exp, doc, NA),
         delta_doc = first(doc) - doc_end_exp, 
         tdelta_doc = first(doc) - last(na.omit(doc)),
         bge = ifelse(delta_doc > 1.5, delta_bc/delta_doc, NA)) %>% 
  fill(doc_end_exp:bge, .direction = "downup") %>% 
  ungroup()
glimpse(doc)
```

# Treatment Averages

```{r}
averages <- doc %>% 
  group_by(Treatment, Timepoint) %>% 
  mutate(ave_toc = mean(PTOC),
         sd_toc = sd(PTOC)) %>% 
  ungroup() %>%
  group_by(Treatment) %>% 
  mutate(ave_bioav_doc = mean(bioav_doc),
         sd_bioav_doc = sd(bioav_doc),
         ave_delta_doc = mean(delta_doc),
         sd_delta_doc = sd(delta_doc),
         ave_tdelta_doc = mean(tdelta_doc),
         sd_tdelta_doc = sd(tdelta_doc),
         ave_bge = mean(bge),
         sd_bge = sd(bge)) %>% 
  ungroup()

glimpse(averages)
```

# TOC Curves

```{r fig.height=7, fig.width=6}
averages %>%
  mutate(Treatment = factor(Treatment, levels = levels)) %>%
  drop_na(ave_toc) %>% 
  ggplot(aes(x = days, y = ave_toc, group = interaction(Experiment, Treatment))) +
  geom_errorbar(aes(ymin = ave_toc - sd_toc, ymax = ave_toc + sd_toc, color = Treatment), width = 0.4) +
  geom_point(aes(fill = Treatment), size = 3, shape = 21, alpha = 0.7) +
  geom_line(aes(color = Treatment), size = 1, alpha = 0.7) +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  labs(x = "Days", y = expression("Total Organic Carbon, µmol C L"^-1), color = "", fill = "") +
  theme_bw() +
  #facet_grid(rows = "Treatment", scales = "free") + 
  guides(color = F)
```

TOC exponentially decreased with time in all three graphs except for Mud. Looking at Ash and GNP, the graph indicates carbon being used more readily than Mud. We can infer that the Ash leachate may contain a higher quality of carbon than that of Mud. What is odd is that growth was not affected by the quality of carbon. 

# Bar Plots

```{r}
bar.data <- averages %>% 
  select(Location, Treatment, ave_bioav_doc:sd_bge) %>% 
  distinct()
```
```{r}
bioav <- bar.data %>%
  ggplot(aes(x = factor(Treatment, levels = levels), y =  ave_bioav_doc, group = Treatment))  + 
  geom_col( color = "black", fill = "white", alpha = 1) +
  geom_errorbar(aes(ymin = ave_bioav_doc - sd_bioav_doc, ymax = ave_bioav_doc + sd_bioav_doc), width = 0.1, size = 0.5) +
  labs(x = "", y = expression(paste("Bioavailable DOC Fraction")), color = "") +
  theme_bw()
```

```{r}
deldoc <- bar.data %>%
  ggplot(aes(x = factor(Treatment, levels = levels), y =  ave_delta_doc, group = Treatment))  +
  geom_col(color = "black", fill = "white", alpha = 1) +
  geom_errorbar(aes(ymin = ave_delta_doc - sd_delta_doc, ymax = ave_delta_doc + sd_delta_doc), width = 0.1, size = 0.5) +
  labs(x = "", y = expression(paste("Delta DOC (to stationary), µmol C L"^-1)), color = "") +
  theme_bw() +
  guides(fill = F)
```

```{r}
total_deldoc <- bar.data %>%
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_tdelta_doc, group = Treatment)) +
  geom_col(color = "black", fill = "white", alpha = 1) +
  geom_errorbar(aes(ymin = ave_tdelta_doc - sd_tdelta_doc, ymax = ave_tdelta_doc + sd_tdelta_doc), width = 0.1, size = 0.5) +
  labs(x = "", y = expression(paste("Delta DOC (total), µmol C L"^-1)), color = "") +
  theme_bw() +
  guides(fill = F) 
```

```{r}
bge <- bar.data %>%
  ggplot(aes(x = factor(Treatment, levels = levels), y =  ave_bge, group = Treatment))  +
  geom_col(color = "black", fill = "white", alpha = 1) +
  geom_errorbar(aes(ymin = ave_bge - sd_bge, ymax = ave_bge + sd_bge), width = 0.1, size = 0.5) +
  labs(x = "", y = expression(paste("Bacterial Growth Efficiency")), color = "") +
  theme_bw() +
  guides(fill = F) 
```

```{r message=FALSE, warning=FALSE}
library(patchwork)
```

```{r fig.height=15, fig.width=8}
deldoc / total_deldoc / bioav / bge + plot_annotation(title = "EEMB 144L 2018 Experiments", tag_levels = "a")
```

```{r}
saveRDS(averages, "~/GitHub EEMB 144L/144l_students/Output_Data/2018_TOC.rds")
```

