---
title: "EEMB 144L 2018 Bacterial Abundance"
author: "Samantha Chen"
date: "11/4/2020"
output: github_document
---
# Intro

In this assignment, I will be showing how **individual bottle** bacterial abundance data from EEMB 144L experiments were processed, QC'd, and analyzed.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(readxl)

excel_sheets("~/GitHub EEMB 144L/144l_students/Input_Data/week3/144L_2018_BactAbund.xlsx")
metadata2018 <- read_excel("~/GitHub EEMB 144L/144l_students/Input_Data/week3/144L_2018_BactAbund.xlsx", sheet = "Metadata")
data2018 <- read_excel("~/GitHub EEMB 144L/144l_students/Input_Data/week3/144L_2018_BactAbund.xlsx", sheet = "Data")

```

```{r}
joined2018 <- left_join(metadata2018, data2018)
```

# Preparing Data

```{r}
cells2018 <- joined2018 %>% 
  mutate(Datetime = ymd_hm(Datetime),
         cells = Cells_ml * 1000) %>% 
  group_by(Experiment, Treatment, Bottle) %>% 
  mutate(interv = interval(first(Datetime), Datetime),
         hours = interv/3600,
         days = hours/24) %>%
  ungroup() %>% 
  select(Experiment:DNA_Sample, hours, days, cells) %>%
  drop_na(cells)
  
```

# Growth Curves
```{r fig.height=10, fig.width=6}
custom.colors <- c("Control" = "#2c7fb8", "Ash Leachate" = "#7fcdbb", "Mud Leachate" = "#756bb1", "Glucose_Nitrate_Phosphate" = "#fc8d62")
levels <- c("Control", "Ash Leachate", "Mud Leachate", "Glucose_Nitrate_Phosphate")

cells2018 %>% 
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>% 
  ggplot(aes(x = days, y = cells, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size = 1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color = "black", shape = 21) +
  geom_text(aes(label = dna), size = 12,  color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("Cells, L"^-1)), fill = "") +
  guides(color = F) +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  facet_grid(rows = "Treatment", scales = "free") +
  theme_bw() 
```
Looking at the growth curves, I can see that both the ash and mud leachate stimulates growth similar to that of the Glucose+Nitrate+Phosphorus. What is interesting is that while the ash leachate and GNP abundance crashed on day 4, the mud leachate did not crash, but does show signs of decline.

# Exponential Growth Curves

```{r}
ln_cells2018 <- cells2018 %>% 
  group_by(Treatment, Bottle) %>% 
  mutate(ln_cells = log(cells),
         diff_ln_cells = ln_cells - lag(ln_cells, default = first(ln_cells))) %>% 
  ungroup()
```

```{r fig.height=6, fig.width=15, message=FALSE, warning=FALSE}
ln_cells2018 %>% 
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>% 
  ggplot(aes(x = days, y = diff_ln_cells, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size = 1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color = "black", shape = 21) +
  geom_text(aes(label = dna), size = 12,  color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("Delta ln cells, L"^-1)), fill = "") +
  guides(color = F) +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  facet_grid(Location~Bottle, scales = "free", drop = F) +
  theme_bw()
```

*Exponential Growth Occurrence*

- Control Bottle A ~0-1 d (T0-T1)
- Control Bottle B ~0-1 d (T0-T1)
- Ash Leachate Bottle C ~0-0.5 d (T0-T1)
- Ash Leachate Bottle D ~0-0.5 d (T0-T1)
- Mud Leachate Bottle E ~0-0.5 d (T0-T1)
- Mud Leachate Bottle F ~0-0.5 d (T0-T1)
- G+N+P Bottle G ~0-1 d (T1-T2)
- G+N+P Bottle H ~0-1 d (T1-T2)

All bottles showed exponential growth from day 0 to day 1.

# Growth rates, doubling times, and ∆cell abundances

```{r}
growth2018 <- ln_cells2018 %>% 
  mutate(start_exp = 0,
         end_exp = ifelse(Bottle %in% c("G", "H"), 2, 1)) %>% 
  group_by(Treatment, Bottle) %>% 
  mutate(ln_cells_start_exp = ifelse(Timepoint == start_exp, ln_cells, NA),
         ln_cells_end_exp = ifelse(Timepoint == end_exp, ln_cells, NA), 
         cells_start_exp = ifelse(Timepoint == start_exp, cells, NA), 
         cells_end_exp = ifelse(Timepoint == end_exp, cells, NA), 
         days_start_exp = ifelse(Timepoint == start_exp, days, NA), 
         days_end_exp = ifelse(Timepoint == end_exp, days, NA)) %>% 
  fill(ln_cells_start_exp:days_end_exp, .direction = "updown") %>% 
  mutate(mew = (ln_cells_end_exp - ln_cells_start_exp)/(days_end_exp - days_start_exp), 
         doubling = log(2)/mew,
         delta_cells = cells_end_exp - first(cells)) %>% 
  ungroup()
         
```

# Bacterial Abundance & Change in Bacterial Abundance --> Carbon Units

```{r}
bactcarbon2018 <- growth2018 %>% 
  mutate(bc = cells * (2.5 * 10^-9), 
         delta_bc = delta_cells * (2.5 * 10^-9))
```

# Treatment Averages

```{r}
averages2018 <- bactcarbon2018 %>%
  group_by(Treatment, Timepoint) %>% 
  mutate(ave_bc = mean(bc),
         sd_bc = sd(bc)) %>% 
  ungroup() %>% 
  group_by(Treatment) %>% 
  mutate(ave_mew = mean(mew),
         sd_mew = sd(mew),
         ave_doubling = mean(doubling),
         sd_doubling = sd(doubling),
         ave_delta_cells = mean(delta_cells),
         sd_delta_cells = sd(delta_cells),
         ave_delta_bc = mean(delta_bc),
         sd_delta_bc = sd(delta_bc),
         ave_lag = mean(days_start_exp),
         sd_lag = sd(days_start_exp)) %>% 
  ungroup()
  
glimpse(averages2018) 
```

# Treatment Average Plots

```{r fig.width=10}
averages2018 %>% 
  ggplot(aes(x = days, y = ave_bc), group = interaction(Experiment, Treatment)) +
  geom_errorbar(aes(ymin = ave_bc - sd_bc, ymax = ave_bc + sd_bc, color = factor(Treatment, levels = levels)), width = 0.1) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size = 1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), color = "black", shape = 21, size = 3) +
  facet_grid(rows = "Treatment", scales = "free") +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  labs(x = "Days", y = expression("Bacterial Carbon, µmol C L"^-1), fill = "", color = "") +
  guides(color = F) +
  theme_bw()
```
Looking at this graph, I can see that Ash Leachate's bacterial carbon curve is very similar to that of GNP's (except for the slight lag between day 0-1). In addition, both the GNP and Ash bacterial carbon abundance decline rapidly between day 3-4, while Mud and control has stayed relatively constant. 

# Barplots

```{r}
bar.data2018 <- averages2018 %>% 
  select(Treatment, ave_mew:sd_lag) %>% 
  distinct()
```

```{r}
mew2018 <- bar.data2018 %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_mew), group = Treatment) +
  geom_col(color = "black", fill = "white") +
  geom_errorbar(aes(ymin = ave_mew - sd_mew, ymax = ave_mew + sd_mew), width = 0.1) +
  labs(x = "", y = expression("µ, d"^-1)) +
  theme_bw()
```

```{r}
doubling2018 <- bar.data2018 %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_doubling), group = Treatment) +
  geom_col(color = "black", fill = "white") +
  geom_errorbar(aes(ymin = ave_doubling - sd_doubling, ymax = ave_doubling + sd_doubling), width = 0.1) +
  labs(x = "", y = expression("Doubling Time, d")) +
  theme_bw()
```
```{r fig.height=8, fig.width=20}
delta_bc2018 <- bar.data2018 %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_delta_bc), group = Treatment) +
  geom_col(color = "black", fill = "white") +
  geom_errorbar(aes(ymin = ave_delta_bc - sd_delta_bc, ymax = ave_delta_bc + sd_delta_bc), width = 0.1) +
  labs(x = "", y = expression("Delta Bacterial Carbon, µmol C L"^-1)) +
  theme_bw()

library(patchwork)
delta_bc2018 + mew2018 + doubling2018 + plot_annotation(tag_levels = "a")
```
 
(a) GNP had the highest delta bacterial carbon telling me that GNP had the most amount of bacterial carbon which means the highest abundance of bacteria. Out of mud and ash, it was ash that had more bacterial carbon therefore ash yielded a higher abundance of bacteria. (b) Ash leachate had the highest specific growth rate while control had the lowest (c) This plot shows an inverse relationship between specific growth rate and doubling time in which the higher the specific growth rate, the shorter amount of time it takes for the bacteria to double. As expected, Ash has the shortest doubling time.


```{r}
saveRDS(averages2018, file = "2018_BactAbund.rds")
```

