---
title: "ACIDD Experiment Bacterial Abundance"
author: "Samantha Chen"
date: "11/3/2020"
output: github_document
---
# Intro

This document shows how **individual bottle** bacterial abundance data from ACIDD experiments were processed, QC'd, and analyzed.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(readxl)
```

#Import Data
```{r}
excel_sheets("~/GitHub EEMB 144L/144l_students/Input_Data/week3/ACIDD_Exp_BactAbund.xlsx")

metadata <- read_excel("~/GitHub EEMB 144L/144l_students/Input_Data/week3/ACIDD_Exp_BactAbund.xlsx", sheet = "Metadata")
glimpse(metadata)
#unique(metadata$Experiment)
#unique(metadata$Location)
#unique(metadata$Bottle)
#unique(metadata$Treatment)

data <- read_excel("~/GitHub EEMB 144L/144l_students/Input_Data/week3/ACIDD_Exp_BactAbund.xlsx", sheet = "Data")
glimpse(data)

joined <- left_join(metadata, data)
names(joined)
#summary(joined)
glimpse(joined)
```

# Prepare Data

Convert date and time column values from characters to dates, add columns with time elapsed for each experiment,convert cells/ml to cells/L, subset data to select only VOI and drop NA's

```{r message=FALSE, warning=FALSE}
cells <- joined %>% 
  mutate(Datetime = ymd_hm(Datetime),
         cells = Cells_ml * 1000,
         sd_cells = Cells_ml_sd * 1000) %>% 
  group_by(Experiment, Treatment, Bottle) %>% 
  mutate(interv = interval(first(Datetime), Datetime), 
         hours = interv/3600,
         days = hours/24) %>% 
  ungroup() %>% 
  select(Experiment:Nutrient_Sample, hours, days, cells, sd_cells) %>% 
  drop_na(cells)

glimpse(cells)
```
#Plot Growth Curves

```{r fig.height=4, fig.width=6,message=FALSE, warning=FALSE}
custom.colors <- c("Control" = "#377EB8", "Ash Leachate" = "#4DAF4A", "Santa Barbara" = "#E41A1C", "San Diego" = "#FF7F00")
levels <- c("Control", "Ash Leachate", "San Diego", "Santa Barbara")

cells %>%
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>% 
  ggplot(aes(x = days, y = cells, group = interaction(Experiment, Treatment, Bottle))) +
  geom_errorbar(aes(ymin = cells - sd_cells, ymax = cells + sd_cells, color = factor(Treatment, levels = levels)), width = 0.1) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size = 1) +
  geom_point(aes(fill = Treatment), size = 3, color = "black", shape = 21) +
  geom_text(aes(label = dna), size = 12, color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("Cells, L"^-1)), fill = "") +
  guides(color = F) +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  facet_grid(rows = "Location", scales = "free") +
  theme_bw()
```

We can calculate:

- Total change in cells from initial condition to end of experiment
- Specific growth rates as the slope of ln(abundance) v. time during exponential growth phase
- Doubling time as ln(2) divided by specific growth rate
- Mean of each of these parameters for each treatment

First we'll need to determine where exponential growth occurs in each of the experiments, if it does. So let's plot ln(abundance) v. time

**NOTE about logarithms in R**

log(x) gives the natural log of x, not log base 10
log10(x) gives the log base 10
log2(x) gives log base 2

```{r fig.height=4, fig.width=10, message=FALSE, warning=FALSE}
ln_cells <- cells %>% 
  group_by(Experiment, Treatment, Bottle) %>% 
  mutate(ln_cells = log(cells),
         diff_ln_cells = ln_cells - lag(ln_cells, default = first(ln_cells))) %>% 
  ungroup()

ln_cells %>%
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>% 
  ggplot(aes(x = days, y = diff_ln_cells, group = interaction(Experiment, Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size = 1) +
  geom_point(aes(fill = Treatment), size = 3, color = "black", shape = 21) +
  geom_text(aes(label = dna), size = 12, color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("Δln cells, L"^-1)), fill = "") +
  guides(color = F) +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  facet_grid(Location~Bottle, scales = "free") +
  theme_bw()
```
This plot makes it a little easier to see, with the data that we have, where exponential growth occurs for each bottle:

- SD Bottle A ~3-5 d (T4-T6)
- SD Bottle B ~4-5 d (T5-T6)
- SD Bottle C ~2-3 d (T3-T4)
- SD Bottle D ~2-3 d (T3-T4)

- SB Bottle A ~0.5-1 d (T1-T2)
- SB Bottle B ~0.5-2 d (T1-T3)
- SB Bottle C ~1-3 d (T2-T4)
- SB Bottle D ~1-3 d (T2-T4)

# Calculate growth rates, doubling time, and Delta cell abd
```{r message=FALSE, warning=FALSE}
growth <- ln_cells %>% 
  mutate(exp_start = ifelse(Experiment == "ASH171" & Bottle == "A", 4, NA),
         exp_start = ifelse(Experiment == "ASH171" & Bottle == "B", 5, exp_start),
         exp_start = ifelse(Experiment == "ASH171" & Bottle %in% c("C", "D"), 3,exp_start),
         exp_start = ifelse(Experiment == "ASH172" & Bottle %in% c("A", "B"), 1, exp_start),
         exp_start = ifelse(Experiment == "ASH172" & Bottle %in% c("C", "D"), 2, exp_start),
         
         exp_end = ifelse(Experiment == "ASH171" & Bottle %in% c("A", "B"), 6, 4),
         exp_end = ifelse(Experiment == "ASH172" & Bottle == "A",2, exp_end),
         exp_end = ifelse(Experiment == "ASH172" & Bottle == "B", 3, exp_end)) %>% 
  group_by(Experiment, Treatment, Bottle) %>% 
  mutate(ln_cells_exp_start = ifelse(Timepoint == exp_start, ln_cells, NA),
         ln_cells_exp_end = ifelse(Timepoint == exp_end, ln_cells, NA),
         cells_exp_start = ifelse(Timepoint == exp_start, cells, NA),
         cells_exp_end = ifelse(Timepoint == exp_end, cells, NA),
         days_exp_start = ifelse(Timepoint == exp_start, days, NA),
         days_exp_end = ifelse(Timepoint == exp_end, days, NA)) %>% 
  fill(ln_cells_exp_start:days_exp_end, .direction = "updown") %>% 
  mutate(mew = (ln_cells_exp_end - ln_cells_exp_start)/(days_exp_end - days_exp_start),
         doubling = log(2)/mew,
         delta_cells = cells_exp_end - first(cells)) %>% 
  ungroup()

```

# Convert bacterial abundance and change in bacterial abundance to carbon units

Apply a carbon conversion factor(CCF) to bacterial abundance (cells L^-1) to generate bacterial carbon(micromol carbon L^-1)

We'll apply the average carbon content of bacterioplankton cells from Coastal Japan (~30 fg C cell^-1) reported by [Fukuda et al., 1998](https://aem.asm.org/content/64/9/3352)

```{r}
bactcarbon <- growth %>% 
  mutate(bc = cells * (2.5 * 10^-9),
         delta_bc = delta_cells * (2.5 * 10^-9))

glimpse(bactcarbon)
```

# Calculate treatment averages

```{r}
averages <- bactcarbon %>% 
  group_by(Experiment, Treatment, Timepoint) %>% 
  mutate(ave_bc = mean(bc),
         sd_bc = sd(bc))
subset <- averages %>% 
  ungroup() %>% 
  group_by(Experiment, Treatment) %>% 
  mutate(ave_mew = mean(mew),
         sd_mew = sd(mew),
         ave_doubling = mean(doubling),
         sd_doubling = sd(doubling),
         ave_delta_cells = mean(delta_cells),
         sd_delta_cells = sd(delta_cells),
         ave_delta_bc = mean(delta_bc),
         sd_delta_bc = sd(delta_bc),
         ave_lag = mean(days_exp_start),
         sd_lag = sd(days_exp_start)) %>% 
  ungroup()
  
glimpse(averages)
```

# Plot treatment averages

```{r fig.height=4, fig.width=6}
averages %>% 
  ggplot(aes(x = days, y =ave_bc), group = interaction(Experiment, Treatment)) +
  geom_errorbar(aes(ymin = ave_bc - sd_bc, ymax = ave_bc +sd_bc, color = factor(Treatment, levels = levels)), width = 0.1) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size = 1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), color = "black", shape = 21, size = 3) +
  facet_grid(rows = "Location", scales = "free") +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  labs(x = "Days", y = expression("Bacterial Carbon, µmol C L"^-1), fill = "", color = "") +
  guides(color = F) +
  theme_bw()
```

### Barplots

```{r}
bar.data <- subset %>% 
  select(Location, Treatment, ave_mew:sd_lag) %>% 
  distinct()
```

```{r}
mew <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_mew), group = interaction(Location, Treatment)) +
  geom_errorbar(aes(ymin = ave_mew - sd_mew, ymax = ave_mew + sd_mew), width = 0.1) +
  geom_col(color = "black", fill = "white") +
  facet_grid(~factor(Location, levels = levels), scales = "free") +
  labs(x = "", y = expression("μ, d"^-1)) +
  theme_bw()
```

```{r}
doubling <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_doubling), group = interaction(Location, Treatment)) +
  geom_errorbar(aes(ymin = ave_doubling - sd_doubling, ymax = ave_doubling + sd_doubling), width = 0.1) +
  geom_col(color = "black", fill = "white") +
  facet_grid(~factor(Location, levels = levels), scales = "free") +
  labs(x = "", y = expression("Doubling Time, d")) +
  theme_bw()
```

```{r}
delta_bc <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_delta_bc), group = interaction(Location, Treatment)) +
  geom_errorbar(aes(ymin = ave_delta_bc - sd_delta_bc, ymax = ave_delta_bc + sd_delta_bc), width = 0.1) +
  geom_col(color = "black", fill = "white") +
  facet_grid(~factor(Location, levels = levels), scales = "free") +
  labs(x = "", y = expression("Δ Bacterial Carbon, μmol C L"^-1)) +
  theme_bw()
```

```{r}
lag <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_lag), group = interaction(Location, Treatment)) +
  geom_errorbar(aes(ymin = ave_lag - sd_lag, ymax = ave_lag + sd_lag), width = 0.1) +
  geom_col(color = "black", fill = "white") +
  facet_grid(~factor(Location, levels = levels), scales = "free") +
  labs(x = "", y = "Lag Phase, days") +
  theme_bw()
```

```{r message=FALSE, warning=FALSE}
#install.packages("patchwork")
library(patchwork)
```

```{r fig.height=6, fig.width=8}
lag + delta_bc + mew + doubling + plot_annotation(tag_levels = "a")
```

# Save Data

```{r}
saveRDS(averages, "~/GitHub EEMB 144L/144l_students/Output_Data/ACIDD_exp_Processed_BactAbund.rds")
```

